Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-03-04 appendix.tex
I prove the second rule is co-regular. The proof can be found in Appendix C.3

This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it. For example, if $\textit{rm}(\textit{empty}) \notin l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$. Similar cases hold for $\textit{pri} \geq \textit{maxPri}(l)$, $\textit{pri} > \textit{maxUPri}(l)$, $\textit{matched}(l)$, $\textit{putInSeq}(l,l_1,\textit{pri})$ (let $l_1$ be sub-sequence of $l$). This completes the proof of this lemma. \qed
\end {proof}

@@

Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$) between each pair of $\textit{cal}(\textit{put},a)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$). The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.

@@

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'$, such that the linearization point of every method event in  $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$. Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into method events of $E_{px}$ that are not in $l'_{\textit{L0}}$. Let $E_{l_t}$ be the set of method events in $l_t$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into method events of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

@@

To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{last}(e) = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{=}$, and the maximal priority of $e$ is unmatched $\textit{put}$ with items in $S$ and matched $\textit{put}$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_2^{>}$, and the maximal priority in $s$ is unmatched $\textit{put}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one unmatched $\textit{put}$, which is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{=}$, and the maximal priority in $e$ is of more than one pair of matched $\textit{put}$ and $\textit{rm}$ with items in $S$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$, where such $\textit{put}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{put}$ with items in $S$, and such $\textit{rm}$ is maximal respect to $<_{\textit{hb}}$ among $\textit{rm}$ with items in $S$.

\item[-] Else, if $\textit{last}(e) = \textit{PQ}_1^{>}$: $e_1$ is generated from $e$ by erasing call and return of one pair of matched $\textit{put}$ and $\textit{rm}$ with maximal priority in $e$.
\end{itemize}

@@

For each such case, we construct a witness automata. We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{rm}(\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{rm}(\textit{empty})$.

@@

Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, it has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a witness automaton.

@@

\section{Co-Regular of Priority Queues}
\label{sec:co-regular of priority queues}

\subsection{Co-Regular of $R_{\textit{pr2}}$}
\label{subsec:co-regular of Rpr2}

By Lemma \ref{lemma:automata for priority queue with single priority}, from now on, it is safe for us to assume that, for each data-differentiated priority queue history $h$ and each $h'$ of projection of $h$, if $h'$ is single-priority, then $\textit{transToQueue}(h')$ is linearizable to queue. One consequence of this assumption is that, in $h'$, there does not exists an item $a$, such that $\textit{rm}(a)$ happens before $\textit{put}(a)$.

By the proof of Lemma \ref{lemma:Rpr1 is co-regular}, with the help of $\mathcal{A}_{\textit{Rpr1-pprr}}, \mathcal{A}_{\textit{Rpr1-prpr}}, \mathcal{A}_{\textit{Rpr1-rppr}}$ and $\mathcal{A}_{\textit{Rpr1-rprp}}$, from now on, we can safely assume that given a data-differentiated history $h$ and operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ in $h$, there does not exists any items $d_1,\ldots,d_m$, such that $x$ is not covered by items $d_1,\ldots,d_m$ and the priority of $d_1,\ldots,d_m$ is less that $x$.

However, above two assumptions are not enough for imply the linearizability to priority queue. One example is shown in \figurename~\ref{fig:history introduct ob order1}. The history $h$ of \figurename~\ref{fig:history introduct ob order1} is not linearizable, even if $h \vert_{1}$ and $h \vert_{2}$ are both linearizable, and either $a$ or $b$ is covered by other items. One explanation is that, since $\textit{put}(b,1)$ happens before $\textit{put}(a,1)$, in choosing linearization points, we should make $\textit{rm}(b)$ before $\textit{rm}(a)$. However, to satisfy the requirements of $R_{\textit{pr2}}$, the linearization points of $\textit{rm}(a)$ can only locates in the interval from $\textit{ret}(\textit{rm},d)$ to $\textit{cal}(\textit{put},e,2)$. But, $\textit{rm}(b)$ is after this entire interval, and we can not locate the linearization point of $\textit{rm}(b)$ in this interval. When there is no such phenomenon, such as the history in \figurename~\ref{fig:history introduct ob order2}, the history becomes linearizable.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER1.pdf}
%\vspace{-10pt}
  \caption{above assumptions do not implies linearizability (1)}
  \label{fig:history introduct ob order1}
\end{figure}

%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER3.pdf}
%\vspace{-10pt}
%  \caption{above assumptions do not implies linearizability (2)}
%  \label{fig:history introduct ob order1}
%\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{PIC-HIS-INTRO-OB-ORDER2.pdf}
%\vspace{-10pt}
  \caption{above assumptions do not implies linearizability (2)}
  \label{fig:history introduct ob order2}
\end{figure}

In the following part of this subsection, we use $<_{\textit{ob}}$ order for the case when we can infer that an item is inserted earlier than another item, and we use gap-point to represent the intervals when it is possible to locate linearizaton points of $\textit{rm}(a)$.

\begin{definition}[gap-point for matched $\textit{put}$ and $\textit{rm}$ operations]\label{def:gap-point for matched put and rm operations}

Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x),\textit{rm}(x)$ of $h$, we say $o$ is a gap-point of $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies that,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] In $\textit{gapPSeq}(h)$, $o$ is after $\textit{call}(\textit{put},x,\textit{pri}_x)$ and $\textit{call}(\textit{rm},x)$, and is before $\textit{ret}(\textit{rm},x)$.
\item[-] For each item $y$ with non-$\textit{pri}_x$ priority in $\textit{gapPSeq}(h)$. If $\textit{put}(y,\_)$ and $\textit{rm}(y)$ is in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$ and $\textit{call}(\textit{rm},x)$ is after $o$; Else if $\textit{put}(y,\_)$ is in $h$ and $\textit{rm}(y)$ is not in $h$, then it can not be the case that $\textit{ret}(\textit{put},y,\_)$ is before $o$;
\end{itemize}

We say that $o$ is a gap-point of priority $\textit{pri}_x$, if we can obtain a sequence $\textit{gapPSeq}(h)$ from $h$ by adding $o$ to some locations between actions of $h$, and satisfies only the latter requirement.
\end{definition}

\begin{restatable}{lemma}{GapEqualsGapPoint}
\label{lemma:gap eqals the existence of gap point}
Given a data-differentiated history $h$ and two operations $\textit{put}(x,\textit{pri}_x)$ and $\textit{rm}(x)$ of $h$. There is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, if and only if there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.
\end{restatable}

@@

To prove the $\textit{only if}$ direction, we need to prove that: if there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, then there is no gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.

We prove this by contradiction. Assume that there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, but there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$. Since there is no gap-point of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, in the time interval starts from $\textit{cal}(\textit{put},x,\textit{pri}_x)$ and $\textit{cal}(\textit{rm},x)$ and ends at $\textit{ret}(\textit{rm},x)$, there is always some item with non-$\textit{pri}_x$ priority in priority queue. Therefore, it is not hard to prove that there is a cycle going through the left-right constraint of $\textit{put}(x,\_)$ and $\textit{rm}(x)$, and by Lemma \ref{lemma:Gap Equals Constraint for Rpr0}, there is no gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$, which contradicts that there is a gap on $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$.

To prove the $\textit{if}$ direction, assume that there is a gap-point $o$ of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$ and let $h_1$ be the sequence in definition of gap-point, which is generated from $h$ by adding $o$ into $h$. Let sequence $h_2$ be a sequence of call, return and operations, such that the projection of $h_2$ into call and return actions is $h$. $h_2$ is generated as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] First, from $h_1$, transform the point $o$ into $\textit{rm}(x)$.

\item[-] Then for all the other items $y$, (1) adding an operation $\textit{put}(y,p)$ at some position between $\textit{call}(\textit{put},y,\_)$ and $\textit{ret}(\textit{put},y)$, and (2) adding an operation $\textit{rm}(y)$ at some position between $\textit{call}(\textit{rm},y)$ and $\textit{ret}(\textit{rm},y)$.

\item[-] For all items $z$ with the same priority as $x$, the positions for adding operations of $z$ into $h$ is arbitrary (but need to satisfy above restriction. In the following we implicitly require this).

\item[-] For all items $z$ whose priority is different from that of $x$,

    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $\textit{ret}(\textit{put},z,\_)$ is before $o$ in $h_1$: Then in $h_2$, $\textit{rm}(z,\_)$ is just after $\textit{call}(\textit{rm},z)$.

    \item[-] If $\textit{put}(z,\_) <_{\textit{hb}} \textit{rm}(z)$, and $o$ is before $\textit{ret}(\textit{put},z,\_)$ $h_1$: Then in $h_2$, $\textit{put}(z,\_)$ is just before $\textit{ret}(\textit{put},z)$.

    \item[-] Otherwise, we can see that $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$. Then we add $o_1=\textit{put}(z,\_)$ and $o_2=\textit{rm}(z)$ in the time interval of interleaving of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ overlaps in $h$, make sure $o_2$ next to $o_1$, and make sure either both of them is before $\textit{rm}(x)$ or both of them is after $\textit{rm}(x)$ in $h_2$.
    \end{itemize}
\end{itemize}

According to the definition of gap-point, it is not hard to see that we can construct such $h_2$. Let $l$ be the projection of $h_2$ into operations, and $l'$ be the projection of $l$ into $x$ and all items that has priorities not equals that of $x$. Let $l'= u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$. It is easy to see that $h \sqsubseteq l$, and $u \cdot v$ contains matched $\textit{put}$ and $\textit{rm}$ operations. Let $L$, $M$ and $R$ be the set of operations in $u$, $v$ and $w$, respectively. Then it is obvious that $L$, $M$ and $R$ is a gap of $\textit{put}(x,\_)$ and $\textit{rm}(x)$ in $h$. \qed
\end {proof}

Given a data-differentiated history $h$ and two items $a,b$, we say that $a <_{\textit{ob}} b$, if the priority of $a$ and $b$ is the same, and one of the following cases holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Type $A$: $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ in $h$,

\item[-] Type $B$: $\textit{rm}(a)$ happens before $\textit{rm}(b)$ in $h$,

\item[-] Type $C$: $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$ in $h$,
\end{itemize}

Let $<_{\textit{ob}}^*$ be the transitive closure of $<_{\textit{ob}}$. Intuitively, $a <_{\textit{ob}} b$ means that $a$ should be inserted earlier than $b$, and $a <_{\textit{ob}}^* b$ means that we can infer that $a$ should be inserted earlier than $b$ with the help of other items. We can explicitly write $a <_{\textit{ob}}^A b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$. Similarly, we can explicitly write $a <_{\textit{ob}}^B b$ and $a <_{\textit{ob}}^C b$ to mean that $a <_{\textit{ob}} b$ and the reason is that $\textit{put}(a,\_)$ happens before $\textit{put}(b,\_)$ and $\textit{rm}(a)$ happens before $\textit{put}(b,\_)$, respectively.

According to the definition of $<_{\textit{ob}}^*$, if $a <_{\textit{ob}}^* b$, then there exists $c_1,\ldots,c_n$, such that $a <_{\textit{ob}} c_1 <_{\textit{ob}} \ldots <_{\textit{ob}} b$. The following lemma states that, the number of intermediate items $c_i$ is in fact bounded.

\begin{restatable}{lemma}{OBOrderHasBoundedLength}
\label{lemma:ob order has bounded length}
Given a data-differentiated history $h$. Assume that $a <_{\textit{ob}} a_1 <_{\textit{ob}} \ldots <_{\textit{ob}} a_m <_{\textit{ob}} b$, then either $a <_{\textit{ob}} b$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} b$ for some $i$, or $a <_{\textit{ob}} a_i <_{\textit{ob}} a_j <_{\textit{ob}} b$ for some $i,j$.
\end{restatable}

>>>>

This is obvious, since for each conditions in the $\textit{Guard}$ part of the rules of priority queue, if a sequence of sequential executions satisfy it, then its sub-sequence also satisfy it. For example, if $\textit{rm}(\textit{empty}) \notin l = u \cdot v \cdot w$ and let $D_l$ be the set of items of $l$, then for each subset $D' \subseteq D_l$, it is obvious that $\textit{rm}(\textit{empty}) \notin l \vert_{ D' }$. Similar cases hold for $\textit{pri} \geq \textit{maxPri}(l)$, $\textit{pri} > \textit{maxUPri}(l)$, $\textit{matched}(l)$, $\textit{putInSeq}(l,l_1,\textit{pri})$ (let $l_1$ be sub-sequence of $l$). This completes the proof of this lemma. \qed
\end {proof}

@@

Given a data-differentiated execution $e$ and its history $h$, we can obtain a sequence $h'$ from $h$ by adding $\textit{put}(a,p)$ (resp., $\textit{rm}(a)$, $\textit{rm}(\textit{empty})$) between each pair of $\textit{cal}(\textit{put},a,p)$ and $\textit{ret}(\textit{put},a)$ (resp., $\textit{cal}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},a)$, $\textit{cal}(\textit{rm},\textit{empty})$ and $\textit{ret}(\textit{rm},\textit{empty})$). The projection of $h'$ into method events is called linearization of $e$ and $h$, and each method event we add in $h'$ can be considered as a linearization point of the corresponding method event.

@@

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'$, such that the linearization point of every method event in  $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $e_{l'}$. Let $E_{px}$ be the set of $\textit{put}$ with priority $\textit{pri}_x$ in $e \setminus x$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into method events of $E_{px}$ that are not in $l'_{\textit{L0}}$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into method events of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

@@

To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially if $e = \epsilon$. When $e \neq \epsilon$, we generate concurrent execution $e_1$ from $e$ as follows: Since $e \in \textit{proj}(e)$ we know that $e \sqsubseteq \textit{MS}(R_e)$, where $R_e = \textit{last}(e)$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $R_e = \textit{PQ}_3$: $e_1$ is generated from $e$ by erasing call and return of one $\textit{rm}(\textit{empty})$.

\item[-] Else, if $R_e = \textit{PQ}_2^{=}, \textit{PQ}_2^{>}, \textit{PQ}_1^{=}, \textit{PQ}_1^{>}$, and the witness is $x$: $e_1$ is generated from $e$ by erasing call and return of method event of $x$.
\end{itemize}

@@

For each such case, we construct a witness automata. We generate witness automata $\mathcal{A}_{\textit{SinPri}}^1$ for the first case, and it is shown in \figurename~\ref{fig:automata for FIFO-1 in appendix}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri})$, $\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},b),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{rm},b)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^2$ for the second case, and it is shown in \figurename~\ref{fig:automata for FIFO-2}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{cal}(\textit{rm},b) + \textit{ret}(\textit{rm},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^3$ for the third case, and it is shown in \figurename~\ref{fig:automata for FIFO-3}. Here $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{ret}(\textit{put},a), \textit{cal}(\textit{rm},a),\textit{ret}(\textit{rm},a),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{ret}(\textit{rm},b)$, $c_4 = c_3 + \textit{cal}(\textit{rm},b)$, $c_5 = c_1 + \textit{ret}(\textit{rm},b)$, $c_6 = c_5 + \textit{cal}(\textit{rm},b)$.

@@

We generate witness automata $\mathcal{A}_{\textit{SinPri}}^4$ for the forth case, and it is shown in \figurename~\ref{fig:automata for FIFO-4}. Here $c_1 = c + \textit{cal}(\textit{rm},b)$, and $c_2 = c + \textit{ret}(\textit{put},b) + \textit{cal}(\textit{rm},a) + \textit{ret}(\textit{rm},a)$, where $c = \textit{cal}(\textit{put},d,\textit{anyPri}),\textit{ret}(\textit{put},d), \textit{cal}(\textit{rm},d),\textit{ret}(\textit{rm},d),\textit{cal}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$.

@@

Let us begin to represent witness automata that is used for capture the existence of a data-differentiated execution $e$, $e$ has a projection $e'$, $\textit{last}(e') = \textit{PQ}_1^{>}$, and there exists a cycle going through the item with maximal priority in $e'$. By data-independence, we can obtain $e_r$ from $e$ by renaming function, which maps such item to be $b$, maps items that cover it to be $a$, and maps other items into $d$. There are four possible enumeration of call and return actions of $\textit{put}(b)$ and $\textit{rm}(b)$. For each of them, we generate a witness automaton.

@@

\subsection{Proofs and Definitions in Subsection \ref{subsec:co-regular of PQ1Equal}}
\label{sec:appendix proof and definition in section co-regular of PQ1Equal}

Let $\textit{Items}(e,p)$ be the set of items with priority $p$ in execution $e$. The following lemma states a method to choose $\textit{itm}$ of $\textit{PQ}_1^{=}$.

\begin{restatable}{lemma}{MaximalInPBadGPMakePQ1Equal}
\label{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}
Given a data-differentiated execution $e$ with $\textit{last}(e) = \textit{PQ}_1^{=}$. Let $\textit{pri}$ be the maximal priority in $e$. If there exists an item $x$ with priority $\textit{pri}$, such that for each $y \in \textit{Items}(e,\textit{pri})$, (1) $x$ does not $<_{\textit{pb}}$ to $y$, and (2) the right-most gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},y)$. Then $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.
\end{restatable}

@@

Let $o$ be the right-most gap-point of $x$. We locate linearization points of each method event as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Locate the linearization point of $\textit{rm}(x)$ at $o$,

\item[-] If $\textit{put}(x,\textit{pri})$ overlaps with $\textit{rm}(x)$, then locate the linearization point of $\textit{put}(x,\textit{pri})$ just before the linearization point of $\textit{rm}(x)$. Otherwise, $\textit{put}(x,\textit{pri}) <_{\textit{hb}} \textit{rm}(x)$, and we locate the linearization point of $\textit{put}(x,\textit{pri})$ just before its return action.

\item[-] Locate linearization points of method event of each $y \in \textit{Items}(e,\textit{pri})$ (except for $x$) just after the call action of the method event.

\item[-] For item $z$ with priority smaller than $\textit{pri}$. If both $\textit{cal}(\textit{put},z,\_)$ and $\textit{cal}(\textit{rm},z)$ is before $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just after their call actions. If both $\textit{ret}(\textit{put},z)$ and $\textit{ret}(\textit{rm},z)$ (if exists) is after $o$, then locate the linearization points of $\textit{put}(z,\_)$ and $\textit{rm}(z)$ just before their return actions. Otherwise, $x$ is in interval of $z$, which contradicts the definition of gap-point.
\end{itemize}

Let $l$ be the sequence of linearization points constructed above. It is obvious that $e \sqsubseteq l$. Since for each $y \in \textit{Items}(e,\textit{pri})$, $o$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$, we can see that $\textit{rm}(x)$ is after $\textit{put}(y,\textit{pri})$ and $\textit{rm}(y)$ in $l$. It is obvious that $\textit{put}(x,\textit{pri})$ is before $\textit{rm}(x)$ in $l$. Since $x$ does not $<_{\textit{pb}}$ to $y$, we can see that no $\textit{put}(y,\textit{pri})$ happens before $\textit{put}(x,\textit{pri})$. Then it is easy to see that $\textit{put}(x,\textit{pri})$ is after $\textit{put}(y,\textit{pri})$ in $l$. Since $\textit{last}(e) = \textit{PQ}_1^{=}$, all other items in $\textit{Items}(e,\textit{pri})$ has matched $\textit{put}$ and $\textit{rm}$, and it is easy to see that they are all before $\textit{rm}(x)$ in $l$.

For item $z$ with priority smaller than $\textit{pri}$, we can see that there are only two possibilities: (1) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ are both before $\textit{rm}(x)$ in $l$, and (2) $\textit{put}(z,\_)$ and $\textit{rm}(z)$ (if exists) are after before $\textit{rm}(x)$ in $l$. Therefore, before $\textit{rm}(x)$ in $l$, the $\textit{put}$ and $\textit{rm}$ of $z$ are matched.

Therefore, it is easy to see that $l \in \textit{MS}(\textit{PQ}_1^{=})$. \qed
\end {proof}

With Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can prove the following lemma, which states that getting rid of case in \figurename~\ref{fig:introduce pb order} is enough for ensure $\textit{last}(e) = \textit{PQ}_1^{=} \Rightarrow e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.

\PQOneEqualAsPBandGP*

To prove the $\textit{if}$ direction, let $e_{x,y}$ be the execution that is obtained from $e$ by erasing all actions of items that has same priority as $x$, except for actions of $x$ and $y$. It is obvious that $\textit{last}(e_{x,y}) = \textit{PQ}_1^{=}$. Since $y <_{\textit{pb}}^* x$, according to $\textit{PQ}_1^{=}$, we can see that $x$ should be chosen as $\textit{itm}$ in $\textit{PQ}_1^{=}$.

According to Lemma \ref{lemma:Lin Equals Constraint for PQ1Lar} (Here we temporarily forget the existence of $y$), the only possible position for locating linearizaton point of $\textit{rm}(x)$ is at gap-point of $x$. Otherwise, if the linearizaton point of $\textit{rm}(x)$ is chosen at a position that is not a gap-point of $x$, then there exists unmatched method event before $\textit{rm}(x)$ with smaller priority. Since the rightmost gap-point of $x$ is before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},y)$, if we locate linearizaton point of $\textit{rm}(x)$ at gap-point of $x$, then $\textit{rm}(x)$ will be before $\textit{cal}(\textit{put},y,\_)$ or $\textit{cal}(\textit{rm},x)$.

Therefore, for every sequence $l = u \cdot \textit{put}(x,\_) \cdot v \cdot \textit{rm}(x) \cdot w$, if $e_{x,y} \sqsubseteq l$, then either $u \cdot v$ contains some unmatched method events of priority less than priority of $x$, or $w$ contains $\textit{put}(y,\_)$ or $\textit{rm}(y)$. In both cases, $l \notin \textit{MS}(\textit{PQ}_1^{=})$.

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume we already know that for each $x$ and $y$ has maximal priority in $e$, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\_)$ and $\textit{cal}(\textit{rm},x)$. We need to prove that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$. Let $\textit{pri}$ be the maximal priority of $e$. Recall that we already assume that each single-priority execution has FIFO property, and item with larger priority is not covered by items with smaller priority.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $e_{\textit{pri}}$ be the projection of $e$ into operations of priority $\textit{pri}$. Since each single-priority execution has FIFO property, there exists sequence $l_{\textit{pri}}$, such that $e_{\textit{pri}} \sqsubseteq l_{\textit{pri}}$, and when we treat $\textit{put}$ as $\textit{enq}$ and $\textit{rm}$ as $\textit{deq}$, $l_{\textit{pri}}$ belongs to queue.

\item[-] Let $a_1$ be the last inserted item of $l_{\textit{pri}}$.

    Step $1$: Check whether for each $b \in \textit{Items}(e,\textit{pri})$, (1) $a_1$ does not $<_{\textit{pb}}$ to $b$, and (2) the right-most gap-point of $a$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$.

    It is easy to see that $a_1$ is of priority $\textit{pri}$, and $a_1$ does not $<_{\textit{pb}}$ to any $b \in \textit{Items}(e,\textit{pri})$. If for each $b \in \textit{Items}(e,\textit{pri})$, the rightmost gap-point of $a_1$ is after $\textit{cal}(\textit{put},b,\textit{pri})$ and $\textit{cal}(\textit{rm},b)$. Then by Lemma \ref{lemma:maximal in pb and gap-point make a candidate of PQ1Equal}, we can obtain that $e \sqsubseteq \textit{MS}(\textit{PQ}_1^{=})$.

\item[-] Otherwise, there exists $a_2 \in \textit{Items}(e,\textit{pri})$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,\textit{pri})$ or $\textit{cal}(\textit{rm},a_2)$ in $e$. We can see that each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.%, and thus, the right-most gap-point of $a_2$ is after the rightmost gap-point of $a_1$.
    By assumption, we know that $a_2$ does not $<_{\textit{pb}}$ to $a_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If for each item $b \in \textit{Items}(e,\textit{pri})$, $a_2$ does not $<_{\textit{pb}}$ to $b$. Then we go to step $1$ and treat $a_2$ similarly as $a_1$.
    \item[-] Otherwise, there exists $a_3$ with priority $\textit{pri}$ such that $a_2 <_{\textit{pb}}^* a_3$.

    Since $l_{\textit{pri}}$ has FIFO property, it is easy to see that there is no cycle in $<_{\textit{pb}}$ order. It is safe to assume that $a_3$ is maximal in the sense of $<_{\textit{pb}}^*$. Or we can say, there does not exists $a_4$, such that $a_3 <_{\textit{pb}}^* a_4$.
